---
title: "Experimental_Model"
author: "Lasse Hyldig Hansen"
date: "2023-04-27"
output: html_document
---

### Loading packages 

```{r}
pacman::p_load(
  tidyverse,
  brms,
  cmdstanr,
  patchwork
)
```

### Loading data

```{r}
cogsci_clean <- read_csv("cogsci_clean.csv")
cogsci_clean2 = cogsci_clean %>% filter(GroupRating > 0) 
```

### Loading the simple bayes model 

```{r}
file <- file.path("stan/W9_SimpleBayes.stan")
mod_simpleBayes <- cmdstan_model(file, cpp_options = list(stan_threads = TRUE),
                     stanc_options = list("O1"))
```

### Loading weighted bayes model

```{r}
file <- file.path("stan/W9_WB.stan")
mod_wb <- cmdstan_model(file, cpp_options = list(stan_threads = TRUE),
                     stanc_options = list("O1"))
```

### Estimating models and setting up data 

```{r}
data_pre_pandemic <- list(
  N = nrow(cogsci_clean2),
  y = (cogsci_clean2$SecondRating)/9,
  Source1 = (cogsci_clean2$FirstRating)/9,
  Source2 = (cogsci_clean2$GroupRating)/9
)

samples_simple <- mod_simpleBayes$sample(
  data = data_pre_pandemic,
  seed = 123,
  chains = 2,
  parallel_chains = 2,
  threads_per_chain = 2,
  iter_warmup = 1500,
  iter_sampling = 3000,
  refresh = 500
)

samples_weighted <- mod_wb$sample(
  data = data_pre_pandemic,
  seed = 123,
  chains = 2,
  parallel_chains = 2,
  threads_per_chain = 2,
  iter_warmup = 1500,
  iter_sampling = 3000,
  refresh = 500
)
```

### Model Diagnostics 
```{r}
samples_weighted$cmdstan_diagnose()
```

```{r}
samples_simple$cmdstan_diagnose()
```
### Model Sample summaries

```{r}
weighted_sample_summary = samples_weighted$summary()
```

```{r}
simple_sample_summary = samples_simple$summary()
```

### Psis visualizations

```{r}
Loo_weighted = samples_weighted$loo(save_psis = TRUE, cores = 4)
plot(Loo_weighted, main = "PSIS Diagnostics Plot - Weighted Bayes")
```

```{r}
Loo_simple = samples_simple$loo(save_psis = TRUE, cores = 4)
plot(Loo_simple, main = "PSIS Diagnostics Plot - Simple Bayes")
```

### Elpd plots

```{r}
elpd <- tibble(
  n = seq(3080),
  weighted_diff_elpd = 
  Loo_weighted$pointwise[, "elpd_loo"] - 
  Loo_simple$pointwise[, "elpd_loo"],
  simple_diff_elpd = 
  Loo_simple$pointwise[, "elpd_loo"] -
  Loo_weighted$pointwise[, "elpd_loo"])

p1 <- ggplot(elpd, aes(x = n, y = weighted_diff_elpd)) +
  geom_point(alpha = .1) +
  #xlim(.5,1.01) +
  #ylim(-1.5,1.5) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  ggtitle("Weighted Bayes Diff Elpd") +
  theme_bw()

p2 <- ggplot(elpd, aes(x = n, y = simple_diff_elpd)) +
  geom_point(alpha = .1) +
  #xlim(.5,1.01) +
  #ylim(-1.5,1.5) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  ggtitle("Simple Bayes Diff Elpd") +
  theme_bw()

pacman::p_load(patchwork)
p1 + p2
```

### Comparing Loo estimates

```{r}
loo_compare(Loo_simple, Loo_weighted)
```
### Getting Loo model weights

```{r}
loo_model_weights(list(Loo_simple, Loo_weighted))
```
### Trace plots - Simple Bayes


```{r}
draws_df <- as_draws_df(samples_simple$draws())

ggplot(draws_df, aes(.iteration, bias, group = .chain, color = .chain)) +
  geom_line(alpha = 0.5) +
  ggtitle("Simple Bayes") +
  theme_classic()
```
### Prior/Posterior plots Simple Bayes 

```{r}
ggplot(draws_df) +
  geom_density(aes(sd), alpha = 0.6, fill = "lightblue") +
  geom_density(aes(sd_prior), alpha = 0.6, fill = "pink") +
  ggtitle("Simple Bayes") +
  theme_bw()
```


```{r}
ggplot(draws_df) +
  geom_density(aes(bias), alpha = 0.6, fill = "lightblue") +
  geom_density(aes(bias_prior), alpha = 0.6, fill = "pink") +
  ggtitle("Simple Bayes") +
  theme_bw()
```

### Trace plots - Weighted Bayes

```{r}
draws_df <- as_draws_df(samples_weighted$draws())

ggplot(draws_df, aes(.iteration, bias, group = .chain, color = .chain)) +
  geom_line(alpha = 0.5) +
  ggtitle("Weighted Bayes") +
  theme_classic()
```

```{r}
ggplot(draws_df, aes(.iteration, w1, group = .chain, color = .chain)) +
  geom_line(alpha = 0.5) +
  ggtitle("Weighted Bayes") +
  theme_classic()
```

```{r}
ggplot(draws_df, aes(.iteration, w2, group = .chain, color = .chain)) +
  geom_line(alpha = 0.5) +
  ggtitle("Weighted Bayes") +
  theme_classic()
```
### Prior/Posterior Plots weighted Bayes 

```{r}
p1 <- ggplot(draws_df) +
  geom_density(aes(bias), alpha = 0.6, fill = "lightblue") +
  geom_density(aes(bias_prior), alpha = 0.6, fill = "pink") +
  ggtitle("Weighted Bayes") +
  theme_bw()

p2 <- ggplot(draws_df) +
  geom_density(aes(w1), alpha = 0.6, fill = "lightblue") +
  geom_density(aes(w1_prior), alpha = 0.6, fill = "pink") +
  ggtitle("Weighted Bayes") +

  theme_bw()

p3 <- ggplot(draws_df) +
  geom_density(aes(w2), alpha = 0.6, fill = "lightblue") +
  geom_density(aes(w2_prior), alpha = 0.6, fill = "pink") +
  ggtitle("Weighted Bayes") +
  theme_bw()

p1 + p2 + p3
```



